The Unicode Cookbook \linebreak for Linguists

### 2018-07-31 book/chapters/implementation.tex
corrections to chapter 8 implementation

\section{How to install Python and R}
\label{installing-python-and-r}
When one encounters problems installing software, or bugs in programming code, search engines are your friend! Installation problems and incomprehensible error messages have typically been encountered and solved by other users. Try simply copying and pasting the output of an error message into a search engine; the solution is often already somewhere online. We are fans of Stack Exchange\footnote{\url{https://stackexchange.com/}} -- a network of question-and-answer websites -- which are extremely helpful in solving issues regarding software installation, bugs in code, etc.

Searching the web for ``install r and python'' returns numerous tutorials on how to set up your machine for scientific data analysis. Note that there is no single correct setup for a particular computer or operating system. Both Python and R are available for Windows, Mac, and Unix operating systems from the Python and R project websites. Another option is to use a so-called package manager, i.e.\ a software program that allows the user to manage software packages and their dependencies. On Mac, we use Homebrew,\footnote{\url{https://brew.sh/}} a simple-to-install (via the Terminal App) free and open source package management system. Follow the instructions on the Homebrew website and then use Homebrew to install R and Python (as well as other software packages such as Git and Jupyter Notebooks).

Alternatively for R, RStudio\footnote{\url{https://www.rstudio.com/}} provides a free and open source integrated development environment (IDE). This application can be downloaded and installed (for Mac, Windows and Unix) and it includes its own R installation and R libraries package manager. For developing in Python, we recommend the free community version of PyCharm,\footnote{\url{https://www.jetbrains.com/pycharm/}} an IDE which is available for Mac, Windows, and Unix.

@@

<<eval=FALSE, tidy=FALSE, engine='bash'>>=
# get the paths to the R executables in bash
pathT=`Rscript -e 'cat(file.path(find.package("qlcData"), 
  "exec", "tokenize"))'`
pathW=`Rscript -e 'cat(file.path(find.package("qlcData"), 
  "exec", "writeprofile"))'`

@@

@@

To make the functionality even more accessible, we have prepared webapps with 
the \texttt{Shiny} framework for the R functions. The webapps are 
included inside the \texttt{qlcData} package and can be started with the 
helper function (in R): \texttt{launch\_shiny('tokenize')}.

@@

% In closing, using GitHub to share code and data provides a platform for sharing scientific results and it also promotes a means for scientific replicability of results. Moreover, we find that in cases where the scientists are building the to tools for analysis, open repositories and data help to ensure that what you see is what you get.

>>>>

\section{How to install Python and R}
\label{installing-python-and-r}

When one encounters problems installing software, or bugs in programming code, search engines are your friend! Installation problems and incomprehensible error messages have typically been encountered and solved by other users. Try simply copying and pasting the output of an error message into a search engine; the solution is often already somewhere online. We are fans of Stack Exchange\footnote{\url{https://stackexchange.com/}} -- a network of question-and-answer websites -- which are extremely helpful in solving issues regarding software installation, bugs in code, etc.

Searching the web for ``install r and python'' returns numerous tutorials on how to set up your machine for scientific data analysis. Note that there is no single correct setup for a particular computer or operating system. Both Python and R are available for Windows, Mac, and Linux operating systems from the Python and R project websites. Another option is to use a so-called package manager, i.e.\ a software program that allows the user to manage software packages and their dependencies. On Mac, we use Homebrew,\footnote{\url{https://brew.sh/}} a simple-to-install (via the Terminal App) free and open source package management system. Follow the instructions on the Homebrew website and then use Homebrew to install R and Python (as well as other software packages such as Git and Jupyter Notebooks).

Alternatively for R, RStudio\footnote{\url{https://www.rstudio.com/}} provides a free and open source integrated development environment (IDE). This application can be downloaded and installed (for Mac, Windows and Linux) and it includes its own R installation and R libraries package manager. For developing in Python, we recommend the free community version of PyCharm,\footnote{\url{https://www.jetbrains.com/pycharm/}} an IDE which is available for Mac, Windows, and Unix.

@@

<<eval=FALSE, tidy=FALSE, engine='bash'>>=
# get the paths to the R executables in bash
pathT=`Rscript -e 'cat(system.file("exec/tokenize", package="qlcData"))'`
pathW=`Rscript -e 'cat(system.file("exec/writeprofile", package="qlcData"))'`

@@

You can also do this within R by using the following commands, again possible replacing \texttt{/user/local/bin} with a suitable location on your system:

<<eval=FALSE>>=
# link executables from within R
file.symlink(system.file("exec/tokenize", package="qlcData"), "/usr/local/bin")
file.symlink(system.file("exec/writeprofile", package="qlcData"), "/usr/local/bin")
@

@@

% To make the functionality even more accessible, we have prepared webapps with 
% the \texttt{Shiny} framework for the R functions. The webapps are 
% included inside the \texttt{qlcData} package and can be started with the 
% helper function (in R): \texttt{launch\_shiny('tokenize')}.

@@

\section{Closing words}

In closing, we hope that these rather elaborate musings on writing systems, Unicode and the IPA will help readers appreciate the progress that has been made over the last decades, but also acknowledge the many pitfalls that are still lurking under the surface. But mainly we hope that our proposals point towards a way forward in sharing scientific data, interpretations and analyses in a more transparent manner.

GitHub (or any similar platform) provides a platform for sharing scientific results and it also promotes a means for scientific replicability of results. Not only is all our code on Github, we actually used Github to collaboratively write this book and we will continue to use it to make corrections and update the book when it will become necessary.

Moreover, we find that in cases where the scientists are building tools for analysis, open repositories and data help to ensure that what you see is what you get.
